The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes

### 2019-09-21 paper2_JNLS/paper.tex
N
NM

@@

@@

The Hamiltonian on either side of the ADC, $\Ham^{\pm}$, generates a corresponding mixed ensemble, or Gibbs measure given by 
\begin{align}
\Gupdn = Z_{\thupdn}^{-1} \exp(-\thupdn \Hupdn) \delta(\En - 1)
\end{align}
% Note: The PNAS paper uses C_theta^{1}. The Majda, Qi Stat Phys paper uses C_theta sometimes and C_theta^{-1} in other places (page 11). So I'll just pick the one I like better.
Here $\theta = \thupdn$ is the inverse temperature and $Z_{\theta}$ a constant (the partition function) that depends on $\theta$. Each measure $\Gupdn$ induces a corresponding ensemble average $\mean{}_{\pm}$. Throughout this paper we use a bracket for ensemble average and overbar for a time average; i.e.~for any quantity $Q$,
\begin{align}
&\mean{Q} = \int_{\Omega}  Q \, d \Gibbs
&&\mbox{\em ensemble average} \\
&\tavg{Q} = \frac{1}{T_2-T_1} \int_{T_1}^{T_2} Q \, dt	 
&&\mbox{\em time average}
\end{align}

@@

Recall that the abrupt depth change is met by traveling waves at some time $t = T_{ADC}$, and for simplicity we set $T_{ADC} = 0$. The KdV equations govern the evolution and nonlinear dispersion of waves over long timescales, $t \gg \delta^{-2}$. Meanwhile, wave evolution over shorter timescales, $t \ll \delta^{-2}$, is much simpler in that it is simply noninteracting traveling waves with uniform wave speed $c = \sqrt{g \depth}$ (i.e. no nonlinearity and no dispersion). Hence, a short instant after waves pass over the ADC, the waveform does not yet have time to alter significantly, which gives the (deterministic) matching condition
\begin{align}
&u(x,t) \vert_{t=0^-} = u(x,t) \vert_{t=0^+} \qquad \text{for } x \in [-\pi, \pi]
%&&\mbox{\em deterministic matching condition}
\end{align}
%or in Johnson's formulation, it is $\depth^{1/4} \eta$ that matches.

Since $u$ matches for every single trajectory, it must also match in the statistical sense. Moreover, the downstream Hamiltonian $\Hdn$ is a functional of $u$ and hence must also match. This gives the {\em statistical matching condition}
\begin{align}
\label{statmatch}
&\meanup{\Hdn} = \meandn{\Hdn}
%&&\mbox{\em statistical matching condition}
\end{align}
Once $\Hdn$ downstream is determined, it is conserved in the outgoing wave field.

%----------------------------------------------------------------------------%
% Skewness formula
%----------------------------------------------------------------------------%
\subsection{Explicit formula for outgoing skewness}
An experimental observation is that the incoming skewness is small, $\meanup{H_3} \approx 0$. Inserting this approximation into the statistical matching condition \eqref{statmatch} and simplifying gives
\begin{equation}
\label{H3H2}
\frac{\meandn{\Hthree^+}} {\meandn{\Htwo^+} - \meanup{\Htwo^+}} = \frac{C_2}{C_3} \dratdn^2
\end{equation}
%where $\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2}$  indicates the difference between the upstream and downstream values.
To convert to dimensional, i.e.~experimental, values we first note that
\begin{align}
\mean{\Hthree^+} = \frac{\pi}{3} \mean{u^3} = 
\frac{1}{3} \pi^{-1/2} \frac{\mean{\eta^3}}{\etastd^3} = \frac{1}{3} \pi^{-1/2} \skw(\eta)
\end{align}
Second, to convert the downstream surface slope, we note
\begin{align}
\pd{u}{\tilde{x}} = \frac{\lamfac \lamdn}{\pi^{3/2} \etastd} \pd{\eta}{\xi}
\end{align}
where slope with respect to $\xi$ is the same as with respect to dimensional $x$.
Note: I believe it is important here to use $\lamdn$ in reverting to dimensional variables since we are matching $\Hdn$.
Hence, we have
\begin{align}
&\mean{\Htwo^+} = \pi \mean{ \left( \pd{u}{\tilde{x}} \right)^2} = 
\left( \frac{\lamfac \lamdn}{\pi \etastd} \right)^2 \var(\eta_x)
\end{align}
%
Then algebra gives
\begin{equation}
\frac{\meandn{\Hthree}} {\meandn{\Htwo} - \meanup{\Htwo}} = 
\frac{\pi^{3/2} \, \etastd^2} {3 \lamfac^2 \lamdn^2} 
\left( \frac{\skwdn(\eta)} { \vardn(\eta_x) - \varup(\eta_x)} \right)
\end{equation}
Also we have the simplification
\begin{equation}
\frac{C_2}{C_3} = \frac{\pi^{3/2} \delup^2}{9 \lamfac^2 \epsup}
\end{equation}
Hence, \eqref{H3H2} gives the remarkably explicit formula
\begin{equation}
\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^3 =
\frac{1}{3} \left( \frac{\ddn}{\etastd} \right)^3
\end{equation}
In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude and as the square of the depth ratio.
% Note: previously I had D^2 on the RHS. The power of 3 comes from using lamdn consistently in the nondimensionalization.

%----------------------------------------------------------%
% Direct numerical simulations
%----------------------------------------------------------%
\section{Direct numerical simulations}

\section{Sampling Algorithms}

\subsection{Naive acceptance-rejection algorithm}

\subsection{Markov-chain Monte Carlo}

\subsection{Improved acceptance-rejection algorithm}

@@

  % Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.85 \linewidth]{Figs/SkewAmp.pdf}
\caption{
Relationship predicted by the theory and confirmed by experiments.
}
\label{AAA}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

\subsection{Comparison of basic features}

\subsection{New experimental measurements guided by theory}

>>>>

% More careful development of theory and experiments, *in conjunction* with one another.

@@

TO DO: Describe the `representative' experiments in which $\dratdn = 0.24$ and $\epsup = 0.017$.
% These use

@@

The Hamiltonian on either side of the ADC, $\Ham^{\pm}$, generates a corresponding mixed ensemble, or Gibbs measure given by 
\begin{align}
\Gupdn = Z_{\thupdn}^{-1} \exp(-\thupdn \Hupdn) \delta(\En - 1)
\end{align}
% Note: The PNAS paper uses C_theta^{1}. The Majda, Qi Stat Phys paper uses C_theta sometimes and C_theta^{-1} in other places (page 11). So I'll just pick the one I like better.
Here $\theta = \thupdn$ is the inverse temperature and $Z_{\theta}$ a constant (the partition function) that depends on $\theta$. Each measure $\Gupdn$ induces a corresponding ensemble average $\mean{}_{\pm}$. Throughout this paper we use a bracket for ensemble average and overbar for a time average; i.e.~for any quantity $Q$,
\begin{align}
&\mean{Q} = \int_{\Omega}  Q \, d \Gibbs
&&\mbox{\em Ensemble average} \\
&\tavg{Q} = \frac{1}{T_2-T_1} \int_{T_1}^{T_2} Q \, dt	 
&&\mbox{\em Time average}
\end{align}

@@

Recall that the abrupt depth change is met by traveling waves at some time $t = T_{ADC}$, and for simplicity we set $T_{ADC} = 0$. The KdV equations govern the evolution and nonlinear dispersion of waves over long timescales, $t \gg \delta^{-2}$. Meanwhile, wave evolution over shorter timescales, $t \ll \delta^{-2}$, is much simpler in that it is simply noninteracting traveling waves with uniform wave speed $c = \sqrt{g \depth}$ (i.e. no nonlinearity and no dispersion). Hence, a short instant after waves pass over the ADC, the waveform does not yet have time to alter significantly, which gives the (deterministic) matching condition
\begin{align}
u(x,t) \vert_{t=0^-} = u(x,t) \vert_{t=0^+},
\qquad \mbox{\em Deterministic matching condition}
\end{align}
holding for $x \in [-\pi, \pi]$.
%or in Johnson's formulation, it is $\depth^{1/4} \eta$ that matches.

Since $u$ matches for every single trajectory, it must also match in the statistical sense. Moreover, the downstream Hamiltonian $\Hdn$ is a functional of $u$ and hence must also match. This gives the {\em statistical matching condition}
\begin{align}
\label{statmatch}
\meanup{\Hdn} = \meandn{\Hdn}
\qquad \mbox{\em Statistical matching condition}
\end{align}
Once $\Hdn$ downstream is determined, it is conserved in the outgoing wave field.

%----------------------------------------------------------------------------%
% Results
%----------------------------------------------------------------------------%
\section{Results}

With the experimental setup described and the theory outlined, we now present results comparing the two. Unless stated otherwise, all parameters used in the theory are taken directly from their experimental values listed in Table (ref). In particular, we set $\delup = 0.22$ in all tests, and for the `representative' experiments, we set $\dratdn = 0.24$ and $\epsup = 0.17$.

% Later: Plot of transfer function for different lambdas. Shows that the transfer function is relatively insensitive to truncation index, as long as N scales appropriately.

\vsp{5}
%----------------------------------------------------------------------------%
% Skewness formula
%----------------------------------------------------------------------------%
\subsection{Explicit formula for outgoing skewness}
We now examine an explicit formula for the outgoing skewness as originally founded in \cite{majda2019statistical}.
An experimental observation is that the incoming skewness is small, $\meanup{H_3} \approx 0$. Inserting this approximation into the statistical matching condition \eqref{statmatch} and simplifying gives
\begin{equation}
\label{H3H2}
\frac{\meandn{\Hthree^+}} {\meandn{\Htwo^+} - \meanup{\Htwo^+}} = \frac{C_2}{C_3} \dratdn^2
\end{equation}
%where $\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2}$  indicates the difference between the upstream and downstream values.
To convert to dimensional, i.e.~experimental, values we first note that
\begin{align}
\mean{\Hthree^+} = \frac{\pi}{3} \mean{u^3} = 
\frac{1}{3} \pi^{-1/2} \frac{\mean{\eta^3}}{\etastd^3} = \frac{1}{3} \pi^{-1/2} \skw(\eta)
\end{align}
Second, to convert the downstream surface slope, we note
\begin{align}
\pd{u}{\tilde{x}} = \frac{\lamfac \lamdn}{\pi^{3/2} \etastd} \pd{\eta}{\xi}
\end{align}
where slope with respect to $\xi$ is the same as with respect to dimensional $x$.
Note: I believe it is important here to use $\lamdn$ in reverting to dimensional variables since we are matching $\Hdn$.
Hence, we have
\begin{align}
&\mean{\Htwo^+} = \pi \mean{ \left( \pd{u}{\tilde{x}} \right)^2} = 
\left( \frac{\lamfac \lamdn}{\pi \etastd} \right)^2 \var(\eta_x)
\end{align}
%
Then algebra gives
\begin{equation}
\frac{\meandn{\Hthree}} {\meandn{\Htwo} - \meanup{\Htwo}} = 
\frac{\pi^{3/2} \, \etastd^2} {3 \lamfac^2 \lamdn^2} 
\left( \frac{\skwdn(\eta)} { \vardn(\eta_x) - \varup(\eta_x)} \right)
\end{equation}
Also we have the simplification
\begin{equation}
\frac{C_2}{C_3} = \frac{\pi^{3/2} \delup^2}{9 \lamfac^2 \epsup}
\end{equation}
Hence, \eqref{H3H2} gives the remarkably explicit formula
\begin{equation}
\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^3 =
\frac{1}{3} \left( \frac{\ddn}{\etastd} \right)^3
\end{equation}
In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude and as the square of the depth ratio.
% Note: previously I had D^2 on the RHS. The power of 3 comes from using lamdn consistently in the nondimensionalization.

\vsp{10}
%----------------------------------------------------------%
% Direct numerical simulations
%----------------------------------------------------------%
\section{Direct numerical simulations}

\section{Sampling Algorithms}
%\subsection{Naive acceptance-rejection algorithm}
%\subsection{Markov-chain Monte Carlo}
%\subsection{Improved acceptance-rejection algorithm}

@@

% Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.85 \linewidth]{Figs/SkewAmp.pdf}
\caption{
Relationship predicted by the theory and confirmed by experiments.
}
\label{AAA}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
 %New experimental measurements guided by theory
